European or American? A re-examination of the transatlantic influence over the IMF

### 2025-07-14 manuscript/manuscript.qmd
literaturereview_writing1.1
explains for macroeconomic factors as covariates.

---
title: European or American? A re-examination of the transatlantic influence over the IMF
date: last-modified
authors:
  - name: Dianyi Yang
    email: d.yang17@lse.ac.uk
    orcid: 0009-0004-4652-3429
    correspondence: true
format:
  #native: default
  #docx: default
  orcid-pdf:
    keep-tex: true
    number-sections: true
    include-in-header:
      - text: |
          \newcommand{\centerstart}{\begin{center}}
          \newcommand{\centerend}{\end{center}}
          \usepackage{amsthm, thm-restate}
          \newtheorem{hyp}{Hypothesis}
          \newtheorem{subhyp}{Hypothesis}[hyp]
          \renewcommand{\thesubhyp}{\thehyp\alph{subhyp}}
          \usepackage{chngcntr}
          \usepackage{apptools}
          \AtAppendix{\counterwithin{figure}{section}}
abstract: This is an abstract.
keywords: IMF, US hegemony, Europe
execute:
  echo: false
  warning: false
  error: false
  message: false
---

@@

-   IMF chief European [@kahler2001]

>>>>

---
title: European or American? A re-examination of the transatlantic influence over the IMF
date: last-modified
authors:
  - name: Dianyi Yang
    email: d.yang17@lse.ac.uk
    orcid: 0009-0004-4652-3429
    correspondence: true
format:
  #native: default
  #docx: default
  orcid-pdf:
    keep-tex: true
    number-sections: true
    include-in-header:
      - text: |
          \newcommand{\centerstart}{\begin{center}}
          \newcommand{\centerend}{\end{center}}
          \usepackage{amsthm, thm-restate}
          \newtheorem{hyp}{Hypothesis}
          \newtheorem{subhyp}{Hypothesis}[hyp]
          \renewcommand{\thesubhyp}{\thehyp\alph{subhyp}}
          \usepackage{chngcntr}
          \usepackage{apptools}
          \AtAppendix{\counterwithin{figure}{section}}
abstract: This is an abstract.
keywords: IMF, US hegemony, Europe
execute:
  echo: false
  warning: false
  error: false
  message: false
bibliography: references.bib
---

@@

As an international organization formally governed by a system of weighted voting and technocratic procedures, IMF is viewed as an “accountant” [@lutz2019], primarily concerned with the internal consistency of fiscal projections to guarantee timely repayment of its financial assistance. However, empirical studies suggested that countries with low international reserves, high current account deficits, and low GDP growth are more likely to receive IMF funding assistance [@barro2005, @knight1997] , despite such indicators often reflecting a limited capacity to repay. This paradox leads to a widely argued statement, particularly among realists, that international institutions are “basically a reflection of the distribution of power in the world” [@mearsheimer1994]. Generally, all studies on IMF lending outcomes agree that major power impose their influence in the process. For example, studies have shown that countries strategically important to the United States are more likely to receive IMF loans, or to receive larger amounts. Specifically, this applies to countries that align with the U.S. in United Nations General Assembly and United Nations Security Council voting [@lipscy2018, @thacker1999, @dreher2009], are in geopolitical vital regions \[\], or hold significant debts to American commercial banks [@oatley2004]. This is not surprising, as the United States, as the most powerful state, has long projected its influence over international financial institutions—much like it has done with the World Bank [@wade2022, @woods2023, @clark2020]. The question this study brings about is, how this reflection of power is distributed within a given international organization, especially one that is widely perceived to be dominated by a few major powers.

The projection of great power capabilities into international organizations is sometimes uneven, because of strategic preferences, the perceived inefficiency of institutional engagement, or other possible reasons. Existing studies provide substantial evidence of U.S. influence over IMF lending decisions, they leave a crucial question open-the comparison between the U.S. and other major powers, particularly the European countries. IMF is perceived to be an institution dominated by a few major power, but the relative weight of American versus European remains ambiguous. The organization is traditionally chaired by an European national[@kahler2001],
